---
title: "descriptives"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(ggcorrplot)
```

```{r}
df <- read_excel(here::here('data/HEBOnewworksheet5.21.xlsx'))

# replace 999 with NA missing value
df[df == 999] <- NA
```

Number of rows and columns in dataset
```{r}
dim(df)
```

Column names of dataset
```{r}
names(df)
```


Create correlation matrix of variables (dropping the "Patient" and "Etiology" columns)
```{r}
corr <- round(cor(df[, 3:ncol(df)], use = 'pairwise.complete.obs'), 2)
```

Correlation p-values

```{r}
p.mat <- cor_pmat(df[, 3:ncol(df)])
```


Plot correlation Matrix
```{r, fig.width=8, fig.height=8}
ggcorrplot(corr, p.mat = p.mat, hc.order = TRUE,
           insig = "blank", tl.cex = 7, tl.srt = 90)
```

This plot shows the significant pairwise correlations within the data.
The white boxes mean that the pairwise correlation is insignificant.
The remaining boxes show significant pairwise correlations between the variables.

What this shows is there are variables that are correlated.
Correlated variables should not be placed into the model together, so the predictors/independent variables need to be discussed.

For example, depression is correlated with ssri (makes sense, since ssri's are prescribed to those with depression)
MELD scores are a combination of bilirubin, sodium, INR, and creatinine values.
