---
title: "Bivariate Statistics"
author: "Daniel Chen"
date: ""
output: 
  html_document: 
    number_sections: yes
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)

df <- readr::read_csv(here::here('data/working/cleaned_hebo.csv'))
```

# ESSLunch

Comparing variables against ESSLunch Scores

## Hydrogen Rise

```{r}
fisher.test(df$lunchESS, df$h2_rise)
```

## Hydrogen Rise, Any

```{r}
fisher.test(df$lunchESS, df$h2_rise_any)
```

## Binned ESSLunch (High/Low)

Bin ESSLunch scores into High/Low values

- scores of 0 or 1 are: low (0)
- otherwise: high (1)

### Hydrogen Rise
```{r}
fisher.test(df$lunchESS_high_low, df$h2_rise)
```

### Hidrogen Rise, Any

```{r}
fisher.test(df$lunchESS_high_low, df$h2_rise_any)
```


# Markers

## Hydrogen Rise
```{r}
fisher.test(df$markers_recode, df$h2_rise)
```

## Hydrogen Rise, Any

```{r}
fisher.test(df$markers_recode, df$h2_rise_any)
```

# HE

HE Stages:

- 0: none
- 1: covert
- 2: overt

## Hydrogen Rise
```{r}
fisher.test(df$HEStage, df$h2_rise)
```

## Hydrogen Rise, Any

```{r}
fisher.test(df$HEStage, df$h2_rise_any)
```

## Binned HE Stages (low/high)

- HE stage of 0, 1: low (0)
- HE stage of 2: high (1)

### Hydrogen Rise

```{r}
fisher.test(df$he_stage_recode, df$h2_rise)
```

### Hydrogen Rise, Any
```{r}
fisher.test(df$he_stage_recode, df$h2_rise_any)
```

# Has Hydrogen

```{r}
# df_he_sub <- df[df$h2_rise == "no rise", ]
# 
# h2_sub <- df %>%
#     select(h2_rise_any, starts_with('ht')) %>%
#     gather(key = 'h_time', value = 'h_value', HT90, HT120)
```

## Hydrogen at 90
```{r}
fit <- aov(HT90 ~ h2_rise, data = df)
summary(fit)
TukeyHSD(fit)
```

## Has Hydrogen at 90, Any Hydrogen

```{r}
fit <- aov(HT90 ~ h2_rise_any, data = df)
summary(fit)
TukeyHSD(fit)
```

## Hydrogen at 120
```{r}
fit <- aov(HT120 ~ h2_rise, data = df)
summary(fit)
TukeyHSD(fit)
```

## Hydrogen at 120, Any Hydrogen
```{r}
fit <- aov(HT120 ~ h2_rise_any, data = df)
summary(fit)
TukeyHSD(fit)
```

## Methane at 90
```{r}
fit <- aov(CH4T90 ~ h2_rise, data = df)
summary(fit)
TukeyHSD(fit)
```

## Methane at 90, Any Hydrogen
```{r}
fit <- aov(CH4T90 ~ h2_rise_any, data = df)
summary(fit)
TukeyHSD(fit)
```

## Methane at 120
```{r}
fit <- aov(CH4T120 ~ h2_rise, data = df)
summary(fit)
TukeyHSD(fit)
```

## Methane at 120, Any Hydrogen
```{r}
fit <- aov(CH4T120 ~ h2_rise_any, data = df)
summary(fit)
TukeyHSD(fit)
```
